Supplemental Material for: The evolution of gene expression in the term placenta of viviparous mammals

### 2016-01-29 manuscript/supplemental_material.Rnw
include collectors for which we have information

\begin{table}
  \begin{tabular}{c c p{3in}}
    \toprule
    Species                       & Experiment Number & Sample Number(s) \\
    \midrule
    \textit{Ateles fusciceps}     &  & \\
    \textit{Bos taurus}           &  & \\
    \textit{Canis familiaris}     &  & \\
    \textit{Dasypus novemcinctus} &  & \\
    \textit{Equus caballus}       &  & SRR850299,  SRR850300 \\
    \textit{Loxodonta africana}   &  & \\
    \textit{Homo sapiens}         & GSE66622 & SRR1850945,
                                               SRR1850946, 
                                               SRR1850951, 
                                               SRR1850952, 
                                               SRR1850953, 
                                               SRR1850954, 
                                               SRR1850957, 
                                               SRR1850958, 
                                               SRR1850965, 
                                               SRR1850966, 
                                               SRR1850975, 
                                               SRR1850976, 
                                               SRR1850983, 
                                               SRR1850984, 
                                               SRR1850987, 
                                               SRR1850988, 
                                               SRR1850991, 
                                               SRR1850992, 
                                               SRR1850993, 
                                               SRR1850994, 
                                               SRR1850995, 
                                               SRR1850996, 
                                               SRR1851003, 
                                               SRR1851004, 
                                               SRR1851011, 
                                               SRR1851012 \\
    \textit{Monodelphis domestica} & & \\
    \textit{Mus musculus}         &  & SRR243594,
                                       SRR243595,
                                       SRR243596,
                                       SRR243597,
                                       SRR243598,
                                       SRR243599,
                                       SRR243600,
                                       SRR243601,
                                       SRR243602,
                                       SRR243586,
                                       SRR243587,
                                       SRR243588,
                                       SRR243589,
                                       SRR243590,
                                       SRR243591,
                                       SRR243592,
                                       SRR243593\\
    \textit{Nannospalax galili}   &  & \\
    \textit{Ovis aries}           &  & ERR489144, ERR489145\\
    \textit{Pan paniscus}         &  & \\
    \textit{Spalax carmeli}       &  & \\
    \textit{Sus scrofa}           &           & SRR543893 \\
    \bottomrule
  \end{tabular}
  \caption{Species, experiment and species numbers used in the analyses presented in this paper}
  \label{tab:srr_numbers}
\end{table}

@@

@@

@@

>>>>

\begin{table}
  \begin{tabular}{c c p{3in}}
    \toprule
    Species                       & Collector & Experiment Number & Sample Number(s) \\
    \midrule
    \textit{Ateles fusciceps}     &  & & \\
    \textit{Bos taurus}           & Bob Kreft & & \\
    \textit{Canis familiaris}     &  & & \\
    \textit{Dasypus novemcinctus} &  & & \\
    \textit{Equus caballus}       & \citet{Coleman.ea2010:Structuralannotationofequine} & SRR850299,  SRR850300 \\
    \textit{Loxodonta africana}   & Kurt Benirschke & & \\
    \textit{Homo sapiens}         & Karina Richani, \citet{Hughes.ea2015:Evaluatingintra-andinter-individual} & GSE66622 
                                             & SRR1850945,
                                               SRR1850946, 
                                               SRR1850951, 
                                               SRR1850952, 
                                               SRR1850953, 
                                               SRR1850954, 
                                               SRR1850957, 
                                               SRR1850958, 
                                               SRR1850965, 
                                               SRR1850966, 
                                               SRR1850975, 
                                               SRR1850976, 
                                               SRR1850983, 
                                               SRR1850984, 
                                               SRR1850987, 
                                               SRR1850988, 
                                               SRR1850991, 
                                               SRR1850992, 
                                               SRR1850993, 
                                               SRR1850994, 
                                               SRR1850995, 
                                               SRR1850996, 
                                               SRR1851003, 
                                               SRR1851004, 
                                               SRR1851011, 
                                               SRR1851012 \\
    \textit{Monodelphis domestica} & K. Smith & & \\
    \textit{Mus musculus}         & J. Gonzalez, \cite{Wang.ea2011:surveyfornovelimprinted} & GSE29690 
                                     & SRR243594,
                                       SRR243595,
                                       SRR243596,
                                       SRR243597,
                                       SRR243598,
                                       SRR243599,
                                       SRR243600,
                                       SRR243601,
                                       SRR243602,
                                       SRR243586,
                                       SRR243587,
                                       SRR243588,
                                       SRR243589,
                                       SRR243590,
                                       SRR243591,
                                       SRR243592,
                                       SRR243593\\
    \textit{Nannospalax galili}   &  & GSE68763 & SRR2016467 \\
    \textit{Ovis aries}           & \cite{Archibald.ea2010:sheepgenomereferencesequence} & ERR489144, ERR489145\\
    \textit{Pan paniscus}         &  & \\
    \textit{Spalax carmeli}       &  & GSE68763 & SRR2016467 \\
    \textit{Sus scrofa}           & \cite{Ahanda.ea2012:Predictionofaltered3-} &  SRP014902 & SRR543893 \\
    \bottomrule
  \end{tabular}
  \caption{Species, experiment and species numbers used in the analyses presented in this paper}
  \label{tab:srr_numbers}
\end{table}

@@

\begin{table}
  \footnotesize
<<go.analysis.all.tissue,echo=FALSE,results="asis">>=
go.analysis.wide <- 
    data.table(dcast(data=go.results,
                     formula=GO.ID+type+Term+Ontology~species,
                     value.var="FDR"))
go.analysis.wide <- 
    go.analysis.wide[,`n(FDR ≤ 0.05)`:=apply(go.analysis.wide[,-(1:4),with=FALSE],1,
                                         function(x){sum(as.numeric(as.character(x)) <= 0.05)})]
go.analysis.wide <- 
    go.analysis.wide[,`max(FDR ≤ 0.05)`:=apply(go.analysis.wide[,-(1:4),with=FALSE],1,
                                           function(x){if (!any(is.finite(x) & x<=0.05)) {
                                                           return(NA)
                                                       } else {
                                                           max(x[is.finite(x) & x<=0.05],na.rm=TRUE)
                                                       }})]
go.analysis.wide.table <- 
    go.analysis.wide[!is.na(`n(FDR ≤ 0.05)`) & `n(FDR ≤ 0.05)` >= 12 &
                         grepl("Top 1% No Housekeeping",type),
                     list(GO.ID,Term,Ontology,`n(FDR ≤ 0.05)`,`max(FDR ≤ 0.05)`)]
go.analysis.wide.table <- 
    go.analysis.wide.table[order(Ontology,-`n(FDR ≤ 0.05)`,`max(FDR ≤ 0.05)`)]
setnames(go.analysis.wide.table,"GO.ID","GO ID")

print(xtable(go.analysis.wide.table,
             display=c("d","s","s","s","d","e"),
             align=  c("c","c","c","c","c","c")
             ),
      math.style.exponents=TRUE,
      include.rownames=FALSE,
      booktabs=TRUE,
      floating=FALSE
      )

@ 
\caption{GO terms with significant enrichment ($\mathrm{FDR}\le0.05$)
  in 12 or more of the species analyzed from genes in the top 1\% of
  expression. $n(\mathrm{FDR}\le0.05)$ is the number of species with
  significant enrichment ($\mathrm{FDR}\le 0.05$) in the GO term
  listed. $max(\mathrm{FDR}\le0.05)$ is the maximum FDR value in all
  species with an FDR of less than $0.05$.}
\label{tab:go_terms_all_species}
\end{table}

@@

\begin{figure}
<<sslp1.evolution.tree,echo=FALSE,fig.width=12,fig.height=12,warning=FALSE,message=FALSE>>=

tree.id <-
    genes.to.tree[combined.fpkm[gene_id=="ENSCAFG00000010638" & species=="canis familiaris",gene_id],tree]
gene.tree.table <- trees[[tree.id]]$table
gene.tree.expression <- combined.fpkm[gene_id %in% gene.tree.table[,gene_id]]
gene.tree <- read.tree(text=trees[[tree.id]]$tree.dat)
plot.tree.matrix(read.tree(text=trees[[tree.id]]$tree.dat),gene.tree.table,gene.tree.expression)
@ 
\caption{\genetreeexplanation{\textit{SSLP1}}}
\label{fig:sslp1_evolution_tree}
\end{figure}

@@

\begin{table}
\footnotesize
<<go.analysis,warning=FALSE,message=FALSE,echo=FALSE,results="asis">>=
placenta.go.results <- rbindlist(list(head(n=20,placenta_go_results_mf[classic.fdr < 0.05,]),
                             head(n=20,placenta_go_results_bp[classic.fdr < 0.05,])))
placenta.go.results <- placenta.go.results[,Annotated:=NULL]
setnames(placenta.go.results,"Significant","# Core")
setnames(placenta.go.results,"classic","p value")
setnames(placenta.go.results,"classic.fdr","FDR")
setnames(placenta.go.results,"GO.ID","GO ID")
setnames(placenta.go.results,"Term","GO Term")
print(xtable(placenta.go.results,
             digits=2,
             display=c("d","s","s","d","f","e","e","s")),
      include.rownames=FALSE,
      booktabs=TRUE,
      floating=FALSE
      )
@ 
\caption{\ac{GO} terms most significantly enriched
  (\Sexpr{placenta.go.results[,sum(Ontology=="MF")]} MF,
  \Sexpr{placenta.go.results[,sum(Ontology=="BP")]} BP)
  in the core placenta transcriptome as compared to all genes
  with 1:1 human orthologs in all species. “\# Core” is the
  number of genes in the core transcriptome for each \ac{GO} term;
  “Expected” is the number of genes expected in the core
  transcriptome for each term; “p-value” are the odds of this
  occurring by chance calculated by the Fisher exact test for
  each term; “FDR” is the false discovery rate as estimated by
  the Benjamini and Hochberg procedure \cite{Benjamini1995:bh_procedure}.}
\label{tab:core_placenta_go_terms}
\end{table}
